Convolutional Neural Network to Detect Equivalent Questions in Online Forums \\ DL-IC 2018 Project

### 2018-07-03 paper/paper_final.tex
paper

%%%%%%%%% ABSTRACT
\begin{abstract}
   The ABSTRACT should be self contained and explain what the paper is about. Usually abstracts are no longer than 300 words. You should state what are the main contributions of your work and tempt the reader to continue to read your paper. A good abstract briefly describes your problem, approach, and key results. This document is based on the CVPR submission template, and it has been adapted to submit a technical report of a project of the Deep Learning and Image Classification course.
\end{abstract}

%%%%%%%%% BODY TEXT
%------------------------------------------------------------------------
\section{Introduction}
Natural Language Processing (NLP) is just one of several fields emerging from Artificial Intelligence. In NLP we want the machine to process and analyze large amount of natural language data to accomplish very interesting tasks (such as Sentiment Analysis, Language Identification and many others). Nowadays, these tasks are plenty and many works on them has been already carried out \footnote{Here is a very interesting repository collecting several NLP tasks along with their main references: \url{https://github.com/Kyubyong/nlp_tasks}}. In this paper, we focus on the task of predicting whether two questions from an online forum are semantically equivalent, i.e. they can be answered by the exact same answer. This can be a very useful tool for Question-Answering community sites as they typically want to keep their database as less redundant as possible, so to make the searches of their users both faster and more effective. Furthermore, they want to keep people from wasting their time answering to questions that have been already solved. \\
The main contribution of this work is \cite{Bogdanova2015Detecting}. In their work, Bogdanova et al. propose a simple Convolutional Neural Network (CNN) to detect semantically equivalent questions. The proposed network first transforms words into \emph{word embeddings} and then applies a convolutional layer which generates question-wide vectors representations. Finally, pairs of questions are scored using a similarity function. The most interesting experimental result is the evaluation of in-domain word embeddings versus the word embeddings generated using all of the English Wikipedia. In particular, they shown that domain-specific word embeddings achieve higher performances. \\
The authors of the paper accurately described the architecture of the network, but no implementation information is provided. Hence, we have thought it could have been quite challenging to replicate this paper as the first deep-learning project of our life. In particular, our efforts focused mainly on three things: 
\begin{itemize}
    \item The training of in-domain word embeddings
    \item The replication of the network architecture presented in the paper
    \item The evaluation of the network using both the word embeddings trained by Google on the English Wikipedia \footnote{\url{http://nlp.stanford.edu/data/glove.6B.zip}} and those trained by us on the same dataset used to train the network
\end{itemize}

%------------------------------------------------------------------------
\section{Related work}
In this section you should discuss published work that relates to your project. This is expected to be full of references, meaning that you have read the existing literature and you know what you are working on very well. This is not just a list or works, but you are supposed to cluster papers that use similar approaches and compare them each other using very short sentences. I strongly suggest to check \cite{steinhardt, lipton} blog posts for some good practices in writing a paper. A quote that I like very much is \emph{``Research is spending 6 hours reading 35 papers, so you can write one sentence containing 2 references''} \cite{twit:ref}. Keep it in mind while you are writing!
%------------------------------------------------------------------------
\section{Proposed approach}
This is the core of your paper, where you describe the details of the proposed method for solving the problem that you set up in the introduction. This is the most important section. It has to be clear why the chosen approach is the right thing to do with respect to the possible alternatives. The explanation of the method has to be readable and understandable and it should not raise obvious questions from the reader.

You can divide the section in paragraphs or subsections that can be useful for the presentation of your method. Usually at this point you may want to place equations, figures and tables to clarify what you are explaining.
%-------------------------------------------------------------------------
\subsection{Mathematics}
Please number all of your sections and displayed equations.  It is
important for readers to be able to refer to any particular equation.  Just
because you didn't refer to it in the text does not mean some future reader
might not need to refer to it.  It is cumbersome to have to use
circumlocutions like ``the equation second from the top of page 3 column
1''. 
%-------------------------------------------------------------------------
\subsection{Footnotes}
Please use footnotes\footnote {This is what a footnote looks like.  It
often distracts the reader from the main flow of the argument.} sparingly.
Indeed, try to avoid footnotes altogether and include necessary peripheral
observations in
the text (within parentheses, if you prefer, as in this sentence).  If you
wish to use a footnote, place it at the bottom of the column on the page on
which it is referenced. Use Times 8-point type, single-spaced.
%-------------------------------------------------------------------------
\subsection{References}
List and number all bibliographical references in 9-point Times,
single-spaced, at the end of your paper. When referenced in the text,
enclose the citation number in square brackets, for
example~\cite{Authors14}.  Where appropriate, include the name(s) of
editors of referenced books.
%-------------------------------------------------------------------------
\subsection{Illustrations, graphs, and photographs}
All graphics should be centered.  Please ensure that any point you wish to
make is resolvable in a printed copy of the paper.  Resize fonts in figures
to match the font in the body text, and choose line widths which render
effectively in print.  Many readers (and reviewers), even of an electronic
copy, will choose to print your paper in order to read it.  You cannot
insist that they do otherwise, and therefore must not assume that they can
zoom in to see tiny details on a graphic.

When placing figures in \LaTeX, it's almost always best to use
\verb+\includegraphics+, and to specify the  figure width as a multiple of
the line width as in the example below
{\small\begin{verbatim}
   \usepackage[dvips]{graphicx} ...
   \includegraphics[width=0.8\linewidth]
                   {myfile.eps}
\end{verbatim}
}

%------------------------------------------------------------------------
\section{Experiments}
In this section you validate your method showing the experiments that you performed. The experiments will vary depending on the project, but you might compare with previously published methods, perform an ablation study to determine the impact of various components of your system, experiment with different hyperparameters or architectural choices, use visualization techniques to gain insight into how your model works, discuss common failure modes of your model, etc. You should include graphs, tables, or other figures to illustrate your experimental results. Divide in subsections or paragraphs to help the reader navigate in your paper.

\paragraph{Datasets.}
Describe the data you are working with for your project. Usually you need to explain what type of data is it, how much data are you working with and if you applied any pre-processing, filtering, or other special treatment to use it. Remember that you have to cite each dataset you used in your project if it has been published from someone else. Instead, if you collected it by yourself you have to describe accurately how you gathered (and labeled) your data.

\paragraph{Experiments setup.}
Here you describe all the architectural choices of your model, the hyper-parameters of your model, \eg optimizer, learning rate, momentum, batch size and if you cross-validate on them.

\paragraph{Results and discussion.}
Discuss your results and compare with other methods. You can also perform an \emph{ablation study} on your model switching on and off some components to understand their contributions.

\section{Conclusion} 
Summarize your key results. What have you learned from the project and suggest future extensions or new applications of your ideas.

\begin{figure}[t]
\begin{center}
\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   %\includegraphics[width=0.8\linewidth]{egfigure.eps}
\end{center}
   \caption{Example of caption.  It is set in Roman so that mathematics
   (always set in Roman: $B \sin A = A \sin B$) may be included without an
   ugly clash.}
\label{fig:long}
\label{fig:onecol}
\end{figure}

\begin{figure*}
\begin{center}
\fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
\end{center}
   \caption{Example of a short caption, which should be centered.}
\label{fig:short}
\end{figure*}

%-------------------------------------------------------------------------
\section*{Acknowledgements}
Here you can thank all the people that have been helpful during the project, but are not coauthors. It is important to add this section only in the final version that you send to the professors and \textbf{NOT in the review version} in order to not compromise the double blind review process. We thank Andrea Romanoni for the insightful comments during the review of the manuscript, and all the old and new Gods who stand beside us when the deadline is dark and full of terrors.

%-------------------------------------------------------------------------
\appendix
\section{Supplementary Material} 
Supplementary material is not counted toward your 4-6 page limit and should be submitted as a separate zip file. Your supplementary material might include:
\begin{itemize}
    \item Source code (if your project proposed an algorithm, or code that is relevant and important for your project.).
    \item Cool videos, interactive visualizations, demos, etc.
\end{itemize}

\paragraph{Examples of things to NOT put in your supplementary material}
\begin{itemize}
    \item The entire PyTorch/TensorFlow Github source code.
    \item Any code that is larger than 10 MB.
    \item Model checkpoints.
    \item A computer virus.
\end{itemize}

\begin{table}
\begin{center}
\begin{tabular}{|l|c|}
\hline
Method & Frobnability \\
\hline\hline
Theirs & Frumpy \\
Yours & Frobbly \\
Ours & Makes one's heart Frob\\
\hline
\end{tabular}
\end{center}
\caption{Results.   Ours is better.}
\end{table}

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
   In the context of online forums, although two questions may seem very different in terms of vocabulary, length and syntax, they may eventually ask the same thing. In this work, we try to detect semantically equivalent questions using a simple Convolutional Neural Network (CNN). The proposed CNN generates distributed vector representations \cite{mikolov2013distributed} for pairs of questions and score them using a similarity metric. We evaluate in-domain \emph{word embeddings} versus the ones trained by Google using all the English Wikipedia \footnote{\url{http://nlp.stanford.edu/data/glove.6B.zip}}: in the former case, the experimental results are very close to the paper \cite{bogdanova2015detecting} we are trying to replicate, while in the latter case they are a bit worse. In particular, in-domain word embeddings do not bring any improvement to the performances as they do in the Bogdanova's work.
\end{abstract}

%%%%%%%%% BODY TEXT
%------------------------------------------------------------------------
\section{Introduction}
Natural Language Processing (NLP) is just one of several fields emerging from Artificial Intelligence. In NLP we want the machine to process and analyze large amount of natural language data to accomplish very interesting tasks (such as Sentiment Analysis, Language Identification and many others). Nowadays, these tasks are plenty and many works on them has been already carried out \footnote{Here is a very interesting repository collecting several NLP tasks along with their main references: \\ \url{https://github.com/Kyubyong/nlp_tasks}}. In this paper, we focus on the task of predicting whether two questions from an online forum are semantically equivalent, i.e. they can be answered by the exact same answer. This can be a very useful tool for Question-Answering community sites as they typically want to keep their databases as less redundant as possible, so to make the searches of their users both faster and more effective. Furthermore, they want to keep people from wasting their time in answering already solved questions. \\
The main contribution of this work is \cite{bogdanova2015detecting}. In their work, Bogdanova et al. propose a simple Convolutional Neural Network (CNN) to detect semantically equivalent questions. The proposed network first transforms words into \emph{word embeddings} and then applies a convolutional layer which generates question-wide vectors representations. Finally, pairs of questions are scored using a similarity function. The most interesting experimental result is the evaluation of in-domain word embeddings versus the word embeddings generated using all of the English Wikipedia. In particular, they shown that domain-specific word embeddings achieve higher performances. \\
In general, the results obtained by our implementation are very close to the ones of the replicated paper. The only shortcoming was that the in-domain word embeddings trained by us did not improve the performances as they did in Bogdnova et al.'s work. We think this is due to the fact that, since the paper does not provide any description of the model employed to train the domain-specific word-embedddings, it is very likely that our model is significantly different from the one they used.\\
The authors of the paper accurately described the architecture of the network, but many crucial details of the model are missing in their description. Hence, we have found quite stimulating to replicate this paper filling the holes with our own ideas. In particular, our efforts focused mainly on the following three aspects that were not explicitly treated in the paper: 
    \paragraph{Handling questions of different sizes.}
    Bogdanova et al. explicitly state that the different sizes of questions is one of the main challenges of their work. Nevertheless, they do not specify the way they solved this issue. We, therefore, tried to experiment two different solutions: fixing the length of each question to the \textbf{maximum} and the \textbf{mean} length of the questions.The second approach resulted to yield the greater accuracy.
    \paragraph{The type of convolution.}
    For the implementation of our network we used a 1-dimensional convolution. Indeed, text can be seen as images with the input maps having only one dimension (height, without width) corresponding to the question length and where the number of input maps is nothing but the size of the word embeddings. The type of convolution represents another missing detail of \cite{bogdanova2015detecting}.    
    \paragraph{The training of in-domain word embeddings.}
    The paper does not provide any information about the model used to train in-domain word embeddings. Therefore, we based our work on a very good tutorial \footnote{\url{http://adventuresinmachinelearning.com/word2vec-keras-tutorial/}} on how to build a the \emph{word2vec} model \cite{mikolov2013distributed} in Keras. However, the experimental results concerning this aspect of our work were not very satisfactory.

%------------------------------------------------------------------------
\section{Related work}
Detecting text similarities is a problem which was probably born before deep learning. In \cite{gomaa2013survey}, Gomaa et al. survey existing algorithms for both lexical and semantical text similarity. All these algorithms have nothing to do with Neural Networks, and can therefore be used as baselines to understand how much deep learning has improved the state-of-the-art.\\
In \cite{kim2014convolutional}, Kim proposes a very simple CNN architecture trained on top of pre-trained word vectors. He experimented the same network on 7 different sentence-level classification tasks and, for each task, he compared the obtained results with the state-of-the-art: the CNN approach improved upon the state-of-the-art on 4 out of 7 tasks. Moreover, he proposes a slight modification of the network with 2 channels of word vectors: one that is kept static during the training and one that is fine-tuned via backpropagation. Although not explicitly stated, we think this has been the main contribution of \cite{bogdanova2015detecting} as the network architecture is pretty much the same. The Kim's architecture represents the state-of-the-art on many NLP sentence classification tasks and it is often used as a layer of more complex networks which need to compute sentence-level distributed vectors representation. For example, in \cite{severyn2015learning}, the network is used as a fundamental preprocessing step in a learning-to-rank algorithm. \\
A different but still interesting approach to the problem of text classification, is described in \cite{zhang2015character}. In this article, the authors explore treating text as a kind of raw signal at character level, and applying temporal (one-dimensional) Convolutional Network to it.\\
Different CNN architectures are used for solving a wide range of NLP tasks and many encouraging results have been already achieved. Yih et al. \cite{yih2014semantic} introduced a novel semantic similarity model based on a CNN architecture for the task of answering single-relation factual questions. Hu et al. \cite{hu2014convolutional} proposed a CNN architecture for hierarchical sentence modeling and sentence matching. Convolution has been also used by Dos Santos et al. \cite{dos2014deep} for sentiment analysis of short texts that jointly uses character-level, word-level and sentence-level information.\\
To tackle the problem of sentences with different lengths, Socher et al. \cite{socher2011dynamic}, introduce a novel dynamic pooling layer which computes a fixed-sized representation from the variable-sized matrices. Another approach could be to use Long Short-Term Memory (LSTM), generating fixed-sized states from the input sentences (as done in \cite{tai2015improved}).\\
In \cite{zhou2015c}, a novel model called C-LSTM is presented: this model stack CNN and LSTM in a unified architecture for semantic sentence modeling. In this way, they are able to combine the strengths of both architectures. In particular, they apply CNN to text data and feed consecutive window features directly to LSTM. In this way, their architecture enables LSTM to learn long-range dependencies from higher-order sequential features.
%------------------------------------------------------------------------
\section{Proposed approach}
In this section we present our strategy for detecting semantically equivalent questions. In subsection 3.1, we describe in detail the architecture of the Neural Network. Then, in subsequent subsections, we focus on the decisions we have made about the things left unspecified by Bogdanova et al.
\subsection{Neural Network Architecture}
As detailed in Figure \ref{fig:network}, the input for the network is tokenized text strings of the two questions. In the first step, the CNN transforms words into real valued feature vectors, also known as \emph{word embeddings}. Next, a convolutional layer is used to construct two distributed vector representations $\vec{r}_{q1}$ and $\vec{r}_{q2}$, one for each input question. Finally, the CNN computes a similarity score between $\vec{r}_{q1}$ and $\vec{r}_{q2}$. Pairs of questions with similarity above a threshold are considered duplicates. Each of these step is explained in more details in the following.
\begin{figure}[t]
\begin{center}
%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
\includegraphics[width=0.8\linewidth]{img/network.png}
\end{center}
\caption{The network architecture.}
\label{fig:network}
%\label{fig:onecol}
\end{figure}

\subsubsection{From text to embeddings}
The first layer of the network transforms words into representations that capture syntactic and semantic information about the words. The initialization of this layer requires two strictly correlated objects:
\begin{itemize}
\item \textbf{The vocabulary}: a dictionary $V$ associating each word of the corpora used to train the embeddings to an integer index. In this vocabulary, we must also reserve an index for all the words which are not present in the vocabulary. Best practices suggest to have the index $0$ for the \emph{unknown} words and then having the other indexes ordered by most common words.
\item \textbf{The pre-trained word-embeddings}: a matrix $W_0 \in \mathbb{R}^{|V| \times d}$ where $d$ is the size the word embeddings. In this matrix, the $i$-th row corresponds to the word embedding of the $i$-th word in the vocabulary.
\end{itemize}•
The input of this layer is a 1-dimensional vector $\vec{q}=[w_1, w_2, ... , w_N]$ where the element $w_i$ is an integer number corresponding to the index of the $i$-th word of the question. We initialize the layer with our pre-trained word embeddings \footnote{either the word embeddings trained by us or the those trained by Google}, which is the matrix $W_0$. This layer, for each index $w_i$ of the input vector, retrieves the corresponding vector representation $\vec{r}^{w_i}$ in matrix $W_0$. Hence, the output is a matrix $X_0 \in \mathbb{R}^{N \times d}$.\\
Notice that if we use pre-trained word embeddings, the values of both the $d$ and $|V|$ have been already decided. These two hyper-parameters should have been tuned by people who trained the matrix $W_0$. Therefore, we will discuss them later when we will describe the training of in-domain word embeddings.  Even the question length $N$ is something that will be treated later on.

\subsubsection{The convolution}
We use a convolutional layer to compute the question-wide distributed vector representations $\vec{r}_{q1}$ and $\vec{r}_{q2}$. For each question, the convolutional layer first produces local features around each word in the question. Then, it combines these local features using a sum operation to create a fixed-sized feature vector (representation) for the question.\\
Given a question $q$, the convolutional layer applies a matrix-vector operation to each window of size $k$ of successive windows in $\vec{q}_{emb} = [\vec{r}^{w_1},\vec{r}^{w_2}, ..., \vec{r}^{w_N}]$. Let us define the vector $\vec{z}_n \in \mathbb{R}^{dk}$ as the concatenation of a sequence of $k$ word embeddings, centralized in the $n$-th word:
\begin{center}
$\vec{z}_n = [\vec{r}^{w_{n-(k-1)/2}},..., \vec{r}^{w_n}, ...,\vec{r}^{w_{n+(k-1)/2)}}]^T$
\end{center}•
The convolutional layer computes the $j$-th element of the vector $\vec{r}_{q1} \in \mathbb{R}^{cl_u}$ as follows: 
\begin{equation}
[\vec{r}_{q1}]_j = f \left( \sum_{1<n<N}[f(W^1 z_n + b^1)]_j \right)
\end{equation}•
where $W^1 \in \mathbb{R}^{cl_u \times dk}$ is the weight matrix of the convolutional layer and $f$ is the hyperbolic tangent function. The same matrix is used to extract local features around each word window of the given question. The global fixed-sized feature vector for the question is obtained by using the sum over all word windows.\\
Matrix $W^1$ and vector $b^1$ are parameters to be learned. The number of convolutional units $cl_u$ (which corresponds to the size of the question representation), and the size of the word context window $k$ are hyper-parameters to be tuned.

\subsubsection{Similarity score}
Given $\vec{r}_{q1}$ and $\vec{r}_{q2}$, i.e. the representations for the input pair of questions $(q_1, q_2)$, the last layer of the CNN computes a similarity score between $q_1$ and $q_2$. In our experiments we use the cosine similarity, i.e. the scalar product of the two normalized vectors:
\begin{center}
$s(q_1, q_2) = \frac{\vec{r}_{q1}}{\lVert \vec{r}_{q1} \rVert} \cdot \frac{\vec{r}_{q2}}{\lVert \vec{r}_{q2} \rVert}$
\end{center}•

%-------------------------------------------------------------------------
\subsection{Question of different length}
As starting point let's understand why unequal text length is a problem in NLP. The basic recipe for processing text in neural models is as follows:
\begin{itemize}\itemsep0.1pt
\item Convert text to tokens
\item Convert tokens to token IDs
\item Look up embedding vectors for IDs and construct a 2D tensor that are input to subsequent layers.
\end{itemize}•
Now, most neural network APIs (such as Keras) are “define-and-run” so we cannot change the dimensions of various matrices and tensors during runtime. All the techniques that can be used to tackle this issue force us to introduce new hyper-parameters or work-around in our model which are not fundamental for  the NLP task but are present to only handle unequal text lengths.\\
We decided to try two of these techniques, which we found out to be the best practices:
\begin{itemize}
	\item Fixing the length of all the questions to the \textbf{maximum} length. All the questions that are shorter are padded using the \emph{UNKNOWN} word (index $0$ of the vocabulary).
	\item Fixing the length of all the questions to the \textbf{mean} length. All the questions that are shorter are padded as in the case above; all the questions that are longer are chopped.
\end{itemize}
To keep the architecture of our network as simpler as possible we decided to apply this techniques as a preprocessing phase, without adding any further layer to the network architecture. Therefore, our network accepts as input only questions of the same length.\\
The experimental results show that the \emph{max} approach provides the best performances. This is probably due to the fact that the \emph{mean} approach is very likely to drop important semantic information from the sentence. It is worth mentioning that, in order for the \emph{max} approach to perform at its best, it can be a good choice dropping the outliers, i.e. the questions that are too long with respect to the other observations.
%-------------------------------------------------------------------------
\subsection{Type of convolution}

%-------------------------------------------------------------------------
\subsection{In-domain Word Embeddings}

%------------------------------------------------------------------------
\section{Experiments}

\paragraph{Datasets.}

\paragraph{Experiments setup.}

\paragraph{Results and discussion.}

\section{Conclusion}

%\begin{figure*}[t]
%\begin{center}
%\fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
%\end{center}
%   \caption{Example of a short caption, which should be centered.}
%\label{fig:short}
%\end{figure*}

%-------------------------------------------------------------------------
\section*{Acknowledgements}

%-------------------------------------------------------------------------
\appendix
\section{Supplementary Material}

%\begin{table}[t]
%\begin{center}
%\begin{tabular}{|l|c|}
%\hline
%Method & Frobnability \\
%\hline\hline
%Theirs & Frumpy \\
%Yours & Frobbly \\
%Ours & Makes one's heart Frob\\
%\hline
%\end{tabular}
%\end{center}
%\caption{Results.   Ours is better.}
%\end{table}
